---
title: "SQL & R - Assignment2"
author: "Jonathan Burns"
date: "2023-09-16"
output: html_document
---

**Packages needed**

```{r message=FALSE, warning=FALSE}
library(odbc)
library(DBI)
library(tidyverse)
library(openintro)
library(RSQL)
library(dplyr)
library(RODBC)
library(missForest)##Tried other more advanced imputation but had to move on
library(mice)##Tried other more advanced imputation but had to move on
library(data.table)
library(pander)
```


**Pulling from my SQL Server (Microsoft)** 
*I also tried finding rown means but the zero values presented an issue.*
```{r}

connection <- odbcConnect("jbdb")
view(connection)

df <- sqlQuery(connection, "select * from dbo.movies_table")

df <- df %>% 
    group_by(movie_name) %>% 
    rowwise() %>% 
    mutate(mean_score = mean(shawn, laura, christina, chris, jon))
```

*I then moved onto replace the zeros with NA's. but this also presented an issue with rowmeans*
```{r}

df <- df %>% 
    group_by(movie_name) %>% 
    rowwise() %>% 
    mutate(mean_score = mean(shawn, laura, christina, chris, jon, na.rm = T)) %>% 
    mutate_all(~na_if(., 0))
view(df)
```


*Found a way to see the missing data better*
```{r}
df_numeric <- df %>%
  select(shawn, laura, christina, chris, jon, movie_name)
md.pattern(df_numeric)
```
**Imputation Work** 
*Using this simple imputation which is based off the mean value it generates not the median, I was able to form a educated guess for the NA values present within 'shawn', 'chris', 'christina', and 'jon' columns (laura did not need imputation).*

```{r}
imputed_shawn <- data.frame(
  original = df_numeric$shawn,
  imputed_zero = replace(df_numeric$shawn, is.na(df_numeric$shawn), 0),
  imputed_mean = replace(df_numeric$shawn, is.na(df_numeric$shawn), mean(df_numeric$shawn, na.rm = TRUE)),
  imputed_median = replace(df_numeric$shawn, is.na(df_numeric$shawn), median(df_numeric$shawn, na.rm = TRUE))
)
imputed_shawn

imputed_chris <- data.frame(
  original = df_numeric$chris,
  imputed_zero = replace(df_numeric$chris, is.na(df_numeric$chris), 0),
  imputed_mean = replace(df_numeric$chris, is.na(df_numeric$chris), mean(df_numeric$chris, na.rm = TRUE)),
  imputed_median = replace(df_numeric$chris, is.na(df_numeric$chris), median(df_numeric$chris, na.rm = TRUE))
)
imputed_chris

imputed_christina <- data.frame(
  original = df_numeric$christina,
  imputed_zero = replace(df_numeric$christina, is.na(df_numeric$christina), 0),
  imputed_mean = replace(df_numeric$christina, is.na(df_numeric$christina), mean(df_numeric$christina, na.rm = TRUE)),
  imputed_median = replace(df_numeric$christina, is.na(df_numeric$christina), median(df_numeric$christina, na.rm = TRUE))
)
imputed_christina

imputed_jon <- data.frame(
  original = df_numeric$jon,
  imputed_zero = replace(df_numeric$jon, is.na(df_numeric$jon), 0),
  imputed_mean = replace(df_numeric$jon, is.na(df_numeric$jon), mean(df_numeric$jon, na.rm = TRUE)),
  imputed_median = replace(df_numeric$jon, is.na(df_numeric$jon), median(df_numeric$jon, na.rm = TRUE))
)
imputed_jon

imputed_values <- data.frame(df_numeric$movie_name, df_numeric$laura,imputed_shawn$imputed_mean, imputed_chris$imputed_mean, imputed_christina$imputed_mean, imputed_jon$imputed_mean)

setnames(imputed_values, old = c('df_numeric.movie_name', 'df_numeric.laura', 'imputed_shawn.imputed_mean', 'imputed_chris.imputed_mean', 'imputed_christina.imputed_mean', 'imputed_jon.imputed_mean'), new = c('movie_names', 'laura', 'shawn_imp_mean', 'chris_imp_mean', 'christina_imp_mean', 'jon_imp_mean'))

head(imputed_values)

```
```{r}
pandoc.table(imputed_values, style=  'rmarkdown', split.table = Inf)
```

