---
title: "DATA 607 Week 10 JB"
author: "Jonathan Burns"
date: "2023-11-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(stringr)
library(tidytext)
library(janeaustenr)
library(textdata)
library(gutenbergr)
```



##Example from book:
```{r}
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```


Source:http://saifmohammad.com/WebDocs/Lexicons/NRC-Emotion-Lexicon.zip

Citation Info:
Version: 0.92
Publicly Released: 10 July 2011
Created By: Dr. Saif M. Mohammad, Dr. Peter Turney 
Home Page: http://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm

#Using NRC
```{r}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```


##Using Bing 
```{r}
jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```



##Austen Sentiment plot
```{r}
ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```





##Comparison of the sentiment dictionaries

**Entirety of P&P book text**
```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice
```






##AFINN, BING an NRC sentiments
```{r message=FALSE, warning=FALSE}
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

Plot to compare
```{r}
bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```
The three sentiment dictionaries show overall that there is similar positive and negative trajectory throughout Austen's Pride and Predudice novel. Upon further examination however, NRC is lacking in representiaion of negative values, while AFINN and Bing both represent negative better. This is likely due to the structure of NRC compared to the others and how is defines negative and positives.


##Examining a separate text

1952, 60481, 3015 Gutenburg Book Codes.

I will be using Project Gutenburg and its R package to examine sentiment of a different author, Charlotte Perkins Gilman.

```{r}
(afinn <- get_sentiments("afinn"))
```
```{r}
(bing <- get_sentiments("bing"))
```
```{r}
(nrc <- get_sentiments("nrc"))

unique(nrc$sentiment)
```

Iteration for loop to get the correct names replaced for analysis down the road.
```{r}

C_P_G <- gutenberg_download(c(1952, 60481, 3015)) 


replace_values <- c(3015, 60481, 1952)
replacement_names <- c("The Man-Made World", "In this our world", "The Yellow Wallpaper")

#Need to change the names for a cross analysis later

for (i in seq_along(replace_values)) {
  C_P_G$gutenberg_id[C_P_G$gutenberg_id == replace_values[i]] <- replacement_names[i]
}


```


```{r}
tidy_CPG <- C_P_G %>% 
    unnest_tokens(word, text) %>% 
    anti_join(stop_words)


```
Most common words?

The Man-Made World - 3015
In this our world - 60481
The Yellow Wallpaper - 1952
```{r}
tidy_CPG %>% 
    count(word, sort = T)

```


```{r}
tidy_CPG <- C_P_G %>% 
    rename("book" = "gutenberg_id" ) %>% 
    group_by(book) %>% 
    mutate(
        linenumber = row_number()) %>% 
    ungroup() %>% 
    unnest_tokens(word, text)
 
```



```{r}
CPG_sentiment <- tidy_CPG %>%
  inner_join(bing) %>%
  count(book, index = linenumber %/% 30, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

ggplot(CPG_sentiment, aes(index, sentiment, fill = book)) +
    geom_col(show.legend = FALSE) +
    facet_wrap(~book, ncol = 2, scales = "free_x")

```
We looked at anger how about sadness?
Lets visualize sadness like we did for anger. But try using Afinn instead of NRC:
```{r}
afinn_sad <- tidy_CPG %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 10) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

  
ggplot(afinn_sad, aes(index, sentiment, fill = "method"))+
  geom_col(show.legend = F) +
  facet_wrap(~method, scales = "free_y")


```

Does this change with NRC and Bing?
```{r}
sadness_nrc <- nrc %>%
    filter(sentiment == "sadness")

tidy_CPG %>% 
    inner_join(sadness_nrc) %>% 
    count(word, sort = T)

CPG_sent <- tidy_CPG %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 10, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
  
ggplot(CPG_sent, aes(index, sentiment, fill = book))+
  geom_col(show.legend = F) +
  facet_wrap(~book, scales = "free_y", nrow = 2 )
```




##Conclusion:
The data above further highlights the length difference and how it can effect visualization and understanding of these text mining
packages. Overall, the first iteration of using Bing has provided the best look into the three books from Charlotte Gilman. I wanted to use Gilmans works because The Wallpaper was one of the darkest things ive read, and I wanted to see if these lexicons could pick up on that. Also see if there were any significant trends with her other work. While it isnt apparent in her other work, the analysis definitely shows that her Yellow Wallpaper short story is much darker than the others.

















