---
title: "Week 3 assignment 607"
author: "Jonathan Burns"
date: "2023-09-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r message=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(stringr)
```


**Question 1 code** 
*provide code that identifies the majors that contain either "DATA" or "STATISTICS"*
```{r}
college_majors_df <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/all-ages.csv")
```


```{r}
college_majors_df[str_detect(college_majors_df$Major, "DATA")|str_detect(college_majors_df$Major,"STATISTICS"),]
        
```
**Question 2 code**

```{r}

q2_original = '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"'

str_view(q2_original)
print(q2_original)
```

```{r}
q2_final <- unlist(str_extract_all(q2_original, pattern = "[a-z]+")) 
print(q2_final)


```

**Question 3**

(.)\1\1

*This expression takes a character (.) and then copies that captured number two separate times (\1\1)*

"(.)(.)\\2\\1"

*This expression has two captured groups and reverses the output*

(..)\1

*This expression will take two characters at once and simply repeat them again.*

"(.).\\1.\\1"

*This expression looks to take a single character, match it with the first character and then repeat it using 1\\, match the character and then repeat it again*

"(.)(.)(.).*\\3\\2\\1"

*The expression looks like it is taking three characters (.)(.)(.) and then repeating it in reverse order using a x multiplier*



**Question 4**

Start and end with the same character.

**Part 1** 
```{r}
list <- c('print', 'hello', 'world', 'mom', 'yippy', 'dead', 'harsh', 'essence', 'median', 'mode', 'sandwich', 'standarddeviation', 'eigenvalue', 'pcaanalysis', 'church', 'eleven')

```

Should be 6 words that print after this expression is done running
```{r}
str_view(list, "^(.)(.*\\1$)")


```
**Part 2**
Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)

```{r}
str_view(list, "([A-Za-z][A-Za-z]).*\\1")
```
**Part 3**
Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
*I added a few more words to see if it would take more than 3, for multiple letters and it worked*
```{r}
str_view(list, "([A-Za-z]).*\\1.*\\1")
```









